### 2016-03-03 Projects/DelCo_2014_2015/Reports/FDS_Validation_Report/TeX_Files/FDS_Exp_Description_Report.tex

% ================
% = Introduction =
% ================
\chapter{Introduction}
\label{chap:Introduction}
[Add intro about FDS, purpose of validation, etc.] 
\\~\\
This report describes a series of full-scale fire test experiments that were conducted for the purpose of FDS model validation. Thirteen experiments that followed five different procedures were conducted in two structures designed to replicate typical residential dwellings. During each test, propane was provided to three diffusion flame burners at a measured flow rate. Local measurements of temperature, gas velocity, heat flux, and gas concentrations were collected at various locations throughout the structure while ventilation patterns within the structure were changed by opening and closing certain doors and vents.

@@

\subsection{Construction}
\label{sec:construction}
Each test structure was built on a concrete slab as shown in Fig.~\ref{fig:struct_pics}. The east structure was designed to simulate a single-story residential structure, and the west structure was designed to simulate a two-story residential structure. The first floor of each structure had an outer wall composed of interlocking concrete blocks with equal side lengths of 0.61~m (2~ft). The joints and gaps between the blocks were filled with high temperature insulation.

\begin{figure}[!ht]
	\includegraphics[width=5.25in]{../../Hose_Stream_Report/Figures/Pictures/east_structure}
	\\~\\
	\includegraphics[width=5.25in]{../../Hose_Stream_Report/Figures/Pictures/west_structure}
	\caption[North side of the east and west structures.]{North side of the east (top) and west (bottom) test structures.}
	\label{fig:struct_pics}
\end{figure}

@@

\begin{figure}[!ht]
	\includegraphics[width=6in]{../../Hose_Stream_Report/Figures/Pictures/TJI_support}
	\caption[Ceiling support of the west structure.]{First floor ceiling support of the west structure composed of wood truss joist I-beams. View is of the southeast corner of the structure.}
	\label{fig:TJI}
\end{figure}
\FloatBarrier

The second floor of the west structure was built on the wood ceiling support described above and was connected to the first floor by a stairwell. The second story walls were wood framed with 51~mm (2~in) by 102~mm (4~in) studs set to 0.40~m (16~in) centers. The interior walls were protected by 16~mm (0.625~in) fire rated gypsum board, 16~mm (0.625~in) durarock board, and a second layer of 16~mm (0.625~in) fire rated gypsum board. The exterior walls were protected with 11~mm (0.44~in) oriented strand board and 8~mm (0.31~in) fiber cement lap siding.

\subsection{Layout}
\label{sec:layout}
Dimensioned floor plans of the east and west structures are presented in Figures~\ref{fig:east_dimensioned_plan}~and~\ref{fig:west_dimensioned_plan}, respectively.

\begin{figure}[!ht]
	\includegraphics[width=\columnwidth]{../Figures/Floor_Plans/East_Structure_Dimensioned_Full}
	\caption[Dimensioned floor plan of the east structure.]{Dimensioned floor plan of the east structure. Structure is symmetric across horizontal centerline.}
	\label{fig:east_dimensioned_plan}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=\columnwidth]{..//Figures/Floor_Plans/West_Structure_2nd_Floor_Dimensioned_Full}
	\\~\\
	\includegraphics[width=\columnwidth]{../Figures/Floor_Plans/West_Structure_1st_Floor_Dimensioned_Full}
	\caption[Dimensioned floor plans of the west structure.]{Dimensioned floor plan of the second floor (top) and first floor (bottom) of the west structure.}
	\label{fig:west_dimensioned_plan}
\end{figure}

The interior dimensions of each structure were approximately 6.1~m (20~ft) wide, 11~m (36~ft) long, and 2.4~m (8~ft) high. The stairs connecting the two floors of the west structure started 1.6~m (5.25~ft) off the south wall with a width of 1.2~m (4~ft) off the east wall. The exterior doors of both structures, the stairwell door on the second level of the west structure, and a square roof vent with side length 1.2~m (4~ft) were opened and closed at certain instances during experiments to change the ventilation patterns within the structures.
\FloatBarrier

@@

\subsubsection*{East Structure}
The east structure was instrumented with five bare-bead thermocouple arrays, four bi-directional probe plus solid thermocouple arrays, five total heat flux plus radiometer sensor pairs, and two gas sample inlet pipes at the locations shown in Fig.~\ref{fig:east_instrumentation}.

\begin{figure}[!ht]
	\includegraphics[width=\columnwidth]{../Figures/Floor_Plans/East_Structure_Dimensioned_Instrumentation}
	\caption{Location of instrumentation in the east structure.}
	\label{fig:east_instrumentation}
\end{figure}

@@

\begin{figure}[!ht]
	\includegraphics[width=0.45\columnwidth]{../Figures/Pictures/doorway_BDPs}
	\\~\\
	\includegraphics[width=0.65\columnwidth]{../Figures/Pictures/roof_vent_BDPs.png}
	\caption[Bi-directional probe plus solid thermocouple arrays in east structure.]{Bi-directional probe plus solid thermocouple array in exterior doorway (top) and in roof vent (bottom) of the east structure.}
	\label{fig:BDP_arrays}
\end{figure}

@@

\subsubsection*{West Structure}
The first floor of the west structure was instrumented with three bare-bead thermocouple arrays, three bi-directional probe plus solid thermocouple arrays, and one gas sample inlet pipe, and the second floor was equipped with three bare-bead thermocouple arrays, four bi-directional probe plus solid thermocouple arrays, and one gas sample inlet pipe. The locations of the instrumentation in the west structure is shown in Fig.~\ref{fig:west_instrumentation}.

\begin{figure}[!ht]
	\includegraphics[width=\columnwidth]{../Figures/Floor_Plans/West_Structure_2nd_Floor_Dimensioned_Instrumentation}
	\\~\\
	\includegraphics[width=\columnwidth]{../Figures/Floor_Plans/West_Structure_1st_Floor_Dimensioned_Instrumentation}
	\caption[Location of instrumentation in the west structure.]{Location of instrumentation in the second floor (top) and first floor (bottom) of the west structure.}
	\label{fig:west_instrumentation}
\end{figure}

The thermocouples in the bare-bead thermocouple arrays were spaced in the same manner as the arrays in the east structure. The bi-directional probe and thermocouple arrays contained 8 probes with spacing identical to the arrays at the exterior doors of the east structure. Gas concentrations were pulled through 9.5~mm (0.375~in) diameter stainless steel tubing located 1.2~m (4~ft) above the floor. 
\FloatBarrier

\subsection{Measurement Uncertainty}
\label{sec:Uncertainty}
There are different components of uncertainty in the reported length, mass, temperature, heat flux, gas concentration, differential pressure, and gas velocity measurements. Uncertainties are grouped into two categories according to the method used to estimate them. Type A uncertainties are those which are evaluated by statistical methods, and Type B are those which are evaluated by other means~\cite{Taylor&Kuyatt:1994}. Type B analysis of systematic uncertainties involves estimating the upper (+a) and lower (-a) limits for the quantity in question such that the probability that the value would be in the interval ($\pm$a) is essentially 100\%. After estimating uncertainties by either Type A or B analysis, the uncertainties are combined in quadrature to yield the combined standard uncertainty. Then the combined standard uncertainty is multiplied by a coverage factor of two, which results in the expanded uncertainty with a 95\% confidence interval (2$\sigma$). For some of these components, such as the zero and calibration elements, uncertainties are derived from referenced instrument specifications. For other components, referenced research results and past experience with the instruments provided input in the uncertainty determination.

@@

@@

In this study, total heat flux measurements were made with water-cooled Schimidt-Bolter gauges. The manufacturer reports a $\pm$3\% calibration expanded uncertainty for these devices~\cite{Medtherm:2003}. Results from an international study on total heat flux gauge calibration and response demonstrated that the uncertainty of a Schmidt-Boelter gauge is typically $\pm$8\%~\cite{Pitts:2006}.

@@

@@

@@

\section{East Structure Tests}
Three different test series---Series 4, Series 5, and Series 6---were conducted in the east structure. Each series was composed of three experiments that used an identical procedure to change the ventilation in the structure. The experiments conducted during Series 4 followed a procedure in which each exterior door was opened to change the ventilation pattern. During the experiments in Series 5, each double door was opened and the roof vent was opened and closed. Both the roof vent and west double door were opened during each experiment in Series 6. Figs.~\ref{fig:east_test_4}--\ref{fig:east_test_6} include a schematic and additional details about the procedures used for the experiments during Series 4--6.

@@

\section{West Structure Tests}
Two different test series---Series 22 and Series 24---were conducted in the west structure. Each series was composed of two experiments that used an identical procedure to change the ventilation in the structure. Experiments conducted during Series 22 followed a procedure in which each double door on the first and second floors was opened. Experiments from Series 24 followed a sequence of events that involved opening the interior stairwell door, opening the west double door on both the first and second level, and opening the south side exterior door on the second level of the structure. Figs.~\ref{fig:west_test_22}~and~\ref{fig:west_test_24} include a schematic and additional details about the procedures used for the experiments during Series 22 and 24.

@@

% \begin{landscape}
% \begin{table}[!ht]
% \caption{east structure experimental event times (mm:ss)}
% \begin{tabular}{lcccccccccccccc}
%  \toprule
% \textbf{Test} & 
% \multicolumn{2}{c}{\textbf{\underline{Corner Burner}}} & 
% \multicolumn{2}{c}{\textbf{\underline{Middle Burner}}} & 
% \multicolumn{2}{c}{\textbf{\underline{Center Burner}}} & 
% \multicolumn{2}{c}{\textbf{\underline{W Double Door}}} & 
% \multicolumn{2}{c}{\textbf{\underline{E Double Door}}} & 
% \multicolumn{2}{c}{\textbf{\underline{Single Door}}} & 
% \multicolumn{2}{c}{\textbf{\underline{Roof Vent}}}
% \\
% \textbf{Number} & 
% \textbf{On} & \textbf{Off} & \textbf{On} & \textbf{Off} & \textbf{On} & \textbf{Off} & 
% \textbf{Close} & \textbf{Open} & \textbf{Close} & \textbf{Open} &
% \textbf{Close} & \textbf{Open} & \textbf{Close} & \textbf{Open}
% \\
% \midrule
% % Test 2
% 2 & 0:20 & 16:20 & 3:20 & 14:20 & 6:20 & 12:20 & 
% N/A & 7:20 & N/A & 9:20 & N/A & 10:30 & N/A & N/A \\
% % Test 3
% 3 & 0:20 & 17:20 & 3:20 & 15:20 & 6:20 & 13:20 & 
% N/A & 7:20 & N/A & 9:20 & N/A & 10:20 & N/A & N/A \\
% % Test 4
% 4 & 0:20 & 17:20 & 3:20 & 15:20 & 6:20 & 13:20 & 
% N/A & 7:20 & N/A & 9:20 & N/A & 10:20 & N/A & N/A 
% \\ \multicolumn{15}{c}{ } \\
% % Test 5a
% 5a & 0:15 & 9:55 & 0:30 & 9:35 & 0:45 & 9:15 & 
% N/A & 3:18 & N/A & 6:18 & N/A & N/A & 7:45 & 2:53 \\
% % Test 5b
% 5b & 0:15 & 10:30 & 0:30 & 10:15 & 0:45 & 10:00 & 
% N/A & 3:45 & N/A & 5:15 & N/A & N/A & 8:30 & 2:15 \\
% % Test 5c
% 5c & 0:15 & 9:45 & 0:30 & 9:30 & 0:45 & 9:15 & 
% N/A & 3:45 & N/A & 5:16 & N/A & N/A & 7:28 & 2:15
% \\ \multicolumn{15}{c}{ } \\
% % Test 6a
% 6a & 0:15 & 9:55 & 0:30 & 9:35 & 0:45 & 9:15 & 
% N/A & 3:18 & N/A & N/A & N/A & N/A & N/A & 2:53 \\
% % Test 6b
% 6b & 0:15 & 9:55 & 0:30 & 9:35 & 0:45 & 9:15 & 
% N/A & 3:18 & N/A & N/A & N/A & N/A & N/A & 2:53 \\
% % Test 6c
% 6c & 0:15 & 9:55 & 0:30 & 9:35 & 0:45 & 9:15 & 
% N/A & 3:18 & N/A & N/A & N/A & N/A & N/A & 2:53 \\
% \bottomrule
% \end{tabular}
% \label{table:east_exp_times}
% \end{table}
% \end{landscape}

>>>>

% ================
% = Introduction =
% ================
\chapter{Introduction}
\label{chap:Introduction}

This report describes a series of full-scale fire experiments that were conducted for the purpose of model validation. Thirteen experiments that followed five different procedures were conducted in two structures designed to replicate typical residential dwellings. During each test, propane was provided to three diffusion flame burners at a measured flow rate. Local measurements of temperature, gas velocity, heat flux, and gas concentrations were collected at various locations throughout the structure while ventilation patterns within the structure were changed by opening and closing certain doors and vents.

@@

\subsection{Construction}
\label{sec:construction}
Each test structure was built on a concrete slab as shown in Fig.~\ref{fig:struct_pics}. The East Structure was designed to simulate a single-story residential structure, and the West Structure was designed to simulate a two-story residential structure. The first floor of each structure had an outer wall composed of interlocking concrete blocks with equal side lengths of 0.61~m (2~ft). The joints and gaps between the blocks were filled with high temperature insulation.

\begin{figure}[!ht]
	\includegraphics[width=5.25in]{../../Hose_Stream_Report/Figures/Pictures/east_structure}
	\\~\\
	\includegraphics[width=5.25in]{../../Hose_Stream_Report/Figures/Pictures/west_structure}
	\caption[North side of the east and West Structures.]{North side of the east (top) and west (bottom) test structures.}
	\label{fig:struct_pics}
\end{figure}

@@

\begin{figure}[!ht]
	\includegraphics[width=6in]{../../Hose_Stream_Report/Figures/Pictures/TJI_support}
	\caption[Ceiling support of the West Structure.]{First floor ceiling support of the West Structure composed of wood truss joist I-beams. View is of the southeast corner of the structure.}
	\label{fig:TJI}
\end{figure}
\FloatBarrier

The second floor of the West Structure was built on the wood ceiling support described above and was connected to the first floor by a stairwell. The second story walls were wood framed with 51~mm (2~in) by 102~mm (4~in) studs set to 0.40~m (16~in) centers. The interior walls were protected by 16~mm (0.625~in) fire rated gypsum board, 16~mm (0.625~in) durarock board, and a second layer of 16~mm (0.625~in) fire rated gypsum board. The exterior walls were protected with 11~mm (0.44~in) oriented strand board and 8~mm (0.31~in) fiber cement lap siding.

\subsection{Layout}
\label{sec:layout}
Dimensioned floor plans of the east and West Structures are presented in Figures~\ref{fig:east_dimensioned_plan}~and~\ref{fig:west_dimensioned_plan}, respectively.

\begin{figure}[!ht]
	\includegraphics[width=\columnwidth]{../Figures/Floor_Plans/East_Structure_Dimensioned_Full}
	\caption[Dimensioned floor plan of the East Structure.]{Dimensioned floor plan of the East Structure. Structure is symmetric across horizontal centerline.}
	\label{fig:east_dimensioned_plan}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=\columnwidth]{..//Figures/Floor_Plans/West_Structure_2nd_Floor_Dimensioned_Full}
	\\~\\
	\includegraphics[width=\columnwidth]{../Figures/Floor_Plans/West_Structure_1st_Floor_Dimensioned_Full}
	\caption[Dimensioned floor plans of the West Structure.]{Dimensioned floor plan of the second floor (top) and first floor (bottom) of the West Structure.}
	\label{fig:west_dimensioned_plan}
\end{figure}

The interior dimensions of each structure were approximately 6.1~m (20~ft) wide, 11~m (36~ft) long, and 2.4~m (8~ft) high. The stairs connecting the two floors of the West Structure started 1.6~m (5.25~ft) off the south wall with a width of 1.2~m (4~ft) off the east wall. The exterior doors of both structures, the stairwell door on the second level of the West Structure, and a square roof vent with side length 1.2~m (4~ft) were opened and closed at certain instances during experiments to change the ventilation patterns within the structures.
\FloatBarrier

@@

\subsubsection*{East Structure}
The East Structure was instrumented with five bare-bead thermocouple arrays, four bi-directional probe plus solid thermocouple arrays, five total heat flux plus radiometer sensor pairs, and two gas sample inlet pipes at the locations shown in Fig.~\ref{fig:east_instrumentation}.

\begin{figure}[!ht]
	\includegraphics[width=\columnwidth]{../Figures/Floor_Plans/East_Structure_Dimensioned_Instrumentation}
	\caption{Location of instrumentation in the East Structure.}
	\label{fig:east_instrumentation}
\end{figure}

@@

\begin{figure}[!ht]
	\includegraphics[width=0.45\columnwidth]{../Figures/Pictures/doorway_BDPs}
	\\~\\
	\includegraphics[width=0.65\columnwidth]{../Figures/Pictures/roof_vent_BDPs.png}
	\caption[Bi-directional probe plus solid thermocouple arrays in East Structure.]{Bi-directional probe plus solid thermocouple array in exterior doorway (top) and in roof vent (bottom) of the East Structure.}
	\label{fig:BDP_arrays}
\end{figure}

@@

\subsubsection*{West Structure}
The first floor of the West Structure was instrumented with three bare-bead thermocouple arrays, three bi-directional probe plus solid thermocouple arrays, and one gas sample inlet pipe, and the second floor was equipped with three bare-bead thermocouple arrays, four bi-directional probe plus solid thermocouple arrays, and one gas sample inlet pipe. The locations of the instrumentation in the West Structure is shown in Fig.~\ref{fig:west_instrumentation}.

\begin{figure}[!ht]
	\includegraphics[width=\columnwidth]{../Figures/Floor_Plans/West_Structure_2nd_Floor_Dimensioned_Instrumentation}
	\\~\\
	\includegraphics[width=\columnwidth]{../Figures/Floor_Plans/West_Structure_1st_Floor_Dimensioned_Instrumentation}
	\caption[Location of instrumentation in the West Structure.]{Location of instrumentation in the second floor (top) and first floor (bottom) of the West Structure.}
	\label{fig:west_instrumentation}
\end{figure}

The thermocouples in the bare-bead thermocouple arrays were spaced in the same manner as the arrays in the East Structure. The bi-directional probe and thermocouple arrays contained 8 probes with spacing identical to the arrays at the exterior doors of the East Structure. Gas concentrations were pulled through 9.5~mm (0.375~in) diameter stainless steel tubing located 1.2~m (4~ft) above the floor.
\FloatBarrier

\subsection{Measurement Uncertainty}
\label{sec:Uncertainty}

There are different components of uncertainty in the reported length, mass, temperature, heat flux, gas concentration, differential pressure, and gas velocity measurements. Uncertainties are grouped into two categories according to the method used to estimate them. Type A uncertainties are those which are evaluated by statistical methods, and Type B are those which are evaluated by other means~\cite{Taylor&Kuyatt:1994}. Type B analysis of systematic uncertainties involves estimating the upper (+a) and lower (-a) limits for the quantity in question such that the probability that the value would be in the interval ($\pm$a) is essentially 100\%. After estimating uncertainties by either Type A or B analysis, the uncertainties are combined in quadrature to yield the combined standard uncertainty. Then the combined standard uncertainty is multiplied by a coverage factor of two, which results in the expanded uncertainty with a 95\% confidence interval (2$\sigma$). For some of these components, such as the zero and calibration elements, uncertainties are derived from referenced instrument specifications. For other components, referenced research results and past experience with the instruments provided input in the uncertainty determination.

\subsubsection{Compartment Dimensions}

@@

\subsubsection{Thermocouples}

@@

\subsubsection{Heat Flux Gauges}

Total heat flux measurements were made with water-cooled Schimidt-Bolter gauges. The manufacturer reports a $\pm$3\% calibration expanded uncertainty for these devices~\cite{Medtherm:2003}. Results from an international study on total heat flux gauge calibration and response demonstrated that the uncertainty of a Schmidt-Boelter gauge is typically $\pm$8\%~\cite{Pitts:2006}.

\subsubsection{Gas Sampling}

@@

\subsubsection{Pressure Transducers}

@@

\subsubsection{Bi-Directional Probes}

@@

\section{East Structure Tests}
Three different test series---Series 4, Series 5, and Series 6---were conducted in the East Structure. Each series was composed of three experiments that used an identical procedure to change the ventilation in the structure. The experiments conducted during Series 4 followed a procedure in which each exterior door was opened to change the ventilation pattern. During the experiments in Series 5, each double door was opened and the roof vent was opened and closed. Both the roof vent and west double door were opened during each experiment in Series 6. Figs.~\ref{fig:east_test_4}--\ref{fig:east_test_6} include a schematic and additional details about the procedures used for the experiments during Series 4--6.

@@

\section{West Structure Tests}
Two different test series---Series 22 and Series 24---were conducted in the West Structure. Each series was composed of two experiments that used an identical procedure to change the ventilation in the structure. Experiments conducted during Series 22 followed a procedure in which each double door on the first and second floors was opened. Experiments from Series 24 followed a sequence of events that involved opening the interior stairwell door, opening the west double door on both the first and second level, and opening the south side exterior door on the second level of the structure. Figs.~\ref{fig:west_test_22}~and~\ref{fig:west_test_24} include a schematic and additional details about the procedures used for the experiments during Series 22 and 24.

@@

% \begin{landscape}
% \begin{table}[!ht]
% \caption{East Structure experimental event times (mm:ss)}
% \begin{tabular}{lcccccccccccccc}
%  \toprule
% \textbf{Test} &
% \multicolumn{2}{c}{\textbf{\underline{Corner Burner}}} &
% \multicolumn{2}{c}{\textbf{\underline{Middle Burner}}} &
% \multicolumn{2}{c}{\textbf{\underline{Center Burner}}} &
% \multicolumn{2}{c}{\textbf{\underline{W Double Door}}} &
% \multicolumn{2}{c}{\textbf{\underline{E Double Door}}} &
% \multicolumn{2}{c}{\textbf{\underline{Single Door}}} &
% \multicolumn{2}{c}{\textbf{\underline{Roof Vent}}}
% \\
% \textbf{Number} &
% \textbf{On} & \textbf{Off} & \textbf{On} & \textbf{Off} & \textbf{On} & \textbf{Off} &
% \textbf{Close} & \textbf{Open} & \textbf{Close} & \textbf{Open} &
% \textbf{Close} & \textbf{Open} & \textbf{Close} & \textbf{Open}
% \\
% \midrule
% % Test 2
% 2 & 0:20 & 16:20 & 3:20 & 14:20 & 6:20 & 12:20 &
% N/A & 7:20 & N/A & 9:20 & N/A & 10:30 & N/A & N/A \\
% % Test 3
% 3 & 0:20 & 17:20 & 3:20 & 15:20 & 6:20 & 13:20 &
% N/A & 7:20 & N/A & 9:20 & N/A & 10:20 & N/A & N/A \\
% % Test 4
% 4 & 0:20 & 17:20 & 3:20 & 15:20 & 6:20 & 13:20 &
% N/A & 7:20 & N/A & 9:20 & N/A & 10:20 & N/A & N/A
% \\ \multicolumn{15}{c}{ } \\
% % Test 5a
% 5a & 0:15 & 9:55 & 0:30 & 9:35 & 0:45 & 9:15 &
% N/A & 3:18 & N/A & 6:18 & N/A & N/A & 7:45 & 2:53 \\
% % Test 5b
% 5b & 0:15 & 10:30 & 0:30 & 10:15 & 0:45 & 10:00 &
% N/A & 3:45 & N/A & 5:15 & N/A & N/A & 8:30 & 2:15 \\
% % Test 5c
% 5c & 0:15 & 9:45 & 0:30 & 9:30 & 0:45 & 9:15 &
% N/A & 3:45 & N/A & 5:16 & N/A & N/A & 7:28 & 2:15
% \\ \multicolumn{15}{c}{ } \\
% % Test 6a
% 6a & 0:15 & 9:55 & 0:30 & 9:35 & 0:45 & 9:15 &
% N/A & 3:18 & N/A & N/A & N/A & N/A & N/A & 2:53 \\
% % Test 6b
% 6b & 0:15 & 9:55 & 0:30 & 9:35 & 0:45 & 9:15 &
% N/A & 3:18 & N/A & N/A & N/A & N/A & N/A & 2:53 \\
% % Test 6c
% 6c & 0:15 & 9:55 & 0:30 & 9:35 & 0:45 & 9:15 &
% N/A & 3:18 & N/A & N/A & N/A & N/A & N/A & 2:53 \\
% \bottomrule
% \end{tabular}
% \label{table:east_exp_times}
% \end{table}
% \end{landscape}
